---
title: "Long term station data analysis"
author: "Brian Yurk"
date: "12/4/2020"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
require("knitr")
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir=normalizePath("..")) #set working directory to the project directory
```

```{r}
rm(list=ls())
gc()
library(tidyverse)
library(lubridate)
library(ggplot2)
library(viridis)
```

Some functions here. A few only do one thing.

```{r}
compute_durations <- function(tib){ #computes time elapsed between measurements (hours)
  tib <- tib %>% mutate(dt = as.duration(lag(datetime,1) %--% datetime)/dhours(1))
  return(tib)
}

#computes drift potential from drift potential per time and time interval width
#Note: units are not as Fryberger (VU). Instead they are 100*VU*hr
compute_dpt <- function(tib){ 
  tib <- tib %>% mutate(dpt = dp*dt)
  return(tib)
}

#computes modified drift potential by not allowing time intervals to exceed 3 hours. 
#For those that do, the remainder of the time interval is recorded
#Note: units are not as Fryberger (VU). Instead they are 100*VU*hr
compute_dpt2 <- function(tib,dtmax=3){ 
  tib <- tib %>% mutate(dpt2 = dp*pmin(dt,3),lost_dt = dt-pmin(dt,3))
  return(tib)
}

compute_dpt2_rec_comp <- function(tib){
  tib <- tib %>% mutate(dpt2_n_comp = dpt2*cos(WD*pi/180), dpt2_e_comp = dpt2*sin(WD*pi/180))
  return(tib)
}


#applies the previous functions after dropping na wind speeds
compute_dp_stuff <- function(tib,dtmax = 3){ 
  tib <- tib %>% drop_na(WS) %>% compute_durations() %>%
    compute_dpt() %>% compute_dpt2(dtmax) %>% compute_dpt2_rec_comp()
  return(tib)
}

drop_yrs <- function(tib,yrs){ #drops specified years from data
  tib <- tib[!(year(tib$datetime) %in% yrs),]
  return(tib)
}

#yearly totals: drift potential, by direction, by coast; 
#lost time, number of observations, etc
#Note: units are not as Fryberger (VU). Instead they are 100*U*hr
compute_by_yr_stuff_directional <- function(tib){ 
  tib <- tib %>% select(datetime,dt,dp,WD,dpt,dpt2,lost_dt) %>% 
    mutate(yr = year(datetime)) %>% group_by(yr) %>% 
    summarize(dp_m = mean(dp,na.rm=TRUE),dt_m=mean(dt,na.rm=TRUE),dt_sd=sd(dt,na.rm=TRUE),
              dpt_su=sum(dpt,na.rm=TRUE),dpt2_su=sum(dpt2,na.rm=TRUE),
              lost_dt_su=sum(lost_dt,na.rm=TRUE),
              dp_n = sum(dpt2*(WD>=0)*(WD<11.25)+dpt2*(WD>=348.75)*(WD<=360),na.rm=TRUE),
              dp_nne = sum(dpt2*(WD>=11.25)*(WD<33.75),na.rm=TRUE),
              dp_ne = sum(dpt2*(WD>=33.75)*(WD<56.25),na.rm=TRUE),
              dp_ene = sum(dpt2*(WD>=56.25)*(WD<78.75),na.rm=TRUE),
              dp_e = sum(dpt2*(WD>=78.75)*(WD<101.25),na.rm=TRUE),
              dp_ese = sum(dpt2*(WD>=101.25)*(WD<123.75),na.rm=TRUE),
              dp_se = sum(dpt2*(WD>=123.75)*(WD<146.25),na.rm=TRUE),
              dp_sse = sum(dpt2*(WD>=146.25)*(WD<168.75),na.rm=TRUE),
              dp_s = sum(dpt2*(WD>=168.75)*(WD<191.25),na.rm=TRUE),
              dp_ssw = sum(dpt2*(WD>=191.25)*(WD<213.75),na.rm=TRUE),
              dp_sw = sum(dpt2*(WD>=213.75)*(WD<236.25),na.rm=TRUE),
              dp_wsw = sum(dpt2*(WD>=236.25)*(WD<258.75),na.rm=TRUE),
              dp_w = sum(dpt2*(WD>=258.75)*(WD<281.25),na.rm=TRUE),
              dp_wnw = sum(dpt2*(WD>=281.25)*(WD<303.75),na.rm=TRUE),
              dp_nw = sum(dpt2*(WD>=303.75)*(WD<326.25),na.rm=TRUE),
              dp_nnw = sum(dpt2*(WD>=326.25)*(WD<348.75),na.rm=TRUE),
              dp_sc = sum(dpt2*(WD>=0)*(WD<=45)+dpt2*(WD>=315)*(WD<=360),na.rm=TRUE),
              dp_swc = sum(dpt2*(WD>=0)*(WD<=90)+dpt2*(WD==360),na.rm=TRUE),
              dp_wc = sum(dpt2*(WD>=45)*(WD<=135),na.rm=TRUE),
              dp_nwc = sum(dpt2*(WD>=90)*(WD<=180),na.rm=TRUE),
              dp_nc = sum(dpt2*(WD>=135)*(WD<=225),na.rm=TRUE),
              dp_nec = sum(dpt2*(WD>=180)*(WD<=270),na.rm=TRUE),
              dp_ec = sum(dpt2*(WD>=225)*(WD<=315),na.rm=TRUE),
              dp_sec = sum(dpt2*(WD>=270)*(WD<=360)+dpt2*(WD==0),na.rm=TRUE),
              nobs=n())
  return(tib)
}


#yearly totals: drift potential, by direction, by coast; 
#lost time, number of observations, etc
#Note: units ARE as Fryberger (VU). Fryberger used dimensionless dt
#(proportion of total time, in this case over the year). Fryberger describes using
#time as percentage then divided by 100, so a proportion
#rrdp is rotated resultant drift potential. Actual rdd is rotated 180
compute_by_yr_stuff_directional_vu <- function(tib){ 
  tib <- tib %>% select(datetime,dt,dp,WD,dpt,dpt2,dpt2_n_comp,dpt2_e_comp,lost_dt) %>%
    mutate(yr = year(datetime)) %>%
    group_by(yr) %>% 
    summarize(dpt_su=sum(dpt,na.rm=TRUE)/sum(dt,na.rm=TRUE),dpt2_su=sum(dpt2,na.rm=TRUE)/(sum(dt,na.rm=TRUE)-sum(lost_dt,na.rm=TRUE)),
              dt_su=sum(dt,na.rm=TRUE),
              lost_dt_su=sum(lost_dt,na.rm=TRUE),
              dp_n = (sum(dpt2*(WD>=0)*(WD<11.25)+dpt2*(WD>=348.75)*(WD<=360),na.rm=TRUE))/
                (sum(dt,na.rm=TRUE)-sum(lost_dt,na.rm=TRUE)),
              dp_nne = sum(dpt2*(WD>=11.25)*(WD<33.75),na.rm=TRUE)/(sum(dt,na.rm=TRUE)-sum(lost_dt,na.rm=TRUE)),
              dp_ne = sum(dpt2*(WD>=33.75)*(WD<56.25),na.rm=TRUE)/(sum(dt,na.rm=TRUE)-sum(lost_dt,na.rm=TRUE)),
              dp_ene = sum(dpt2*(WD>=56.25)*(WD<78.75),na.rm=TRUE)/(sum(dt,na.rm=TRUE)-sum(lost_dt,na.rm=TRUE)),
              dp_e = sum(dpt2*(WD>=78.75)*(WD<101.25),na.rm=TRUE)/(sum(dt,na.rm=TRUE)-sum(lost_dt,na.rm=TRUE)),
              dp_ese = sum(dpt2*(WD>=101.25)*(WD<123.75),na.rm=TRUE)/(sum(dt,na.rm=TRUE)-sum(lost_dt,na.rm=TRUE)),
              dp_se = sum(dpt2*(WD>=123.75)*(WD<146.25),na.rm=TRUE)/(sum(dt,na.rm=TRUE)-sum(lost_dt,na.rm=TRUE)),
              dp_sse = sum(dpt2*(WD>=146.25)*(WD<168.75),na.rm=TRUE)/(sum(dt,na.rm=TRUE)-sum(lost_dt,na.rm=TRUE)),
              dp_s = sum(dpt2*(WD>=168.75)*(WD<191.25),na.rm=TRUE)/(sum(dt,na.rm=TRUE)-sum(lost_dt,na.rm=TRUE)),
              dp_ssw = sum(dpt2*(WD>=191.25)*(WD<213.75),na.rm=TRUE)/(sum(dt,na.rm=TRUE)-sum(lost_dt,na.rm=TRUE)),
              dp_sw = sum(dpt2*(WD>=213.75)*(WD<236.25),na.rm=TRUE)/(sum(dt,na.rm=TRUE)-sum(lost_dt,na.rm=TRUE)),
              dp_wsw = sum(dpt2*(WD>=236.25)*(WD<258.75),na.rm=TRUE)/(sum(dt,na.rm=TRUE)-sum(lost_dt,na.rm=TRUE)),
              dp_w = sum(dpt2*(WD>=258.75)*(WD<281.25),na.rm=TRUE)/(sum(dt,na.rm=TRUE)-sum(lost_dt,na.rm=TRUE)),
              dp_wnw = sum(dpt2*(WD>=281.25)*(WD<303.75),na.rm=TRUE)/(sum(dt,na.rm=TRUE)-sum(lost_dt,na.rm=TRUE)),
              dp_nw = sum(dpt2*(WD>=303.75)*(WD<326.25),na.rm=TRUE)/(sum(dt,na.rm=TRUE)-sum(lost_dt,na.rm=TRUE)),
              dp_nnw = sum(dpt2*(WD>=326.25)*(WD<348.75),na.rm=TRUE)/(sum(dt,na.rm=TRUE)-sum(lost_dt,na.rm=TRUE)),
              dp_sc = (sum(dpt2*(WD>=0)*(WD<=45)+dpt2*(WD>=315)*(WD<=360),na.rm=TRUE))/
                (sum(dt,na.rm=TRUE)-sum(lost_dt,na.rm=TRUE)),
              dp_swc = (sum(dpt2*(WD>=0)*(WD<=90)+dpt2*(WD==360),na.rm=TRUE))/
                (sum(dt,na.rm=TRUE)-sum(lost_dt,na.rm=TRUE)),
              dp_wc = sum(dpt2*(WD>=45)*(WD<=135),na.rm=TRUE)/(sum(dt,na.rm=TRUE)-sum(lost_dt,na.rm=TRUE)),
              dp_nwc = sum(dpt2*(WD>=90)*(WD<=180),na.rm=TRUE)/(sum(dt,na.rm=TRUE)-sum(lost_dt,na.rm=TRUE)),
              dp_nc = sum(dpt2*(WD>=135)*(WD<=225),na.rm=TRUE)/(sum(dt,na.rm=TRUE)-sum(lost_dt,na.rm=TRUE)),
              dp_nec = sum(dpt2*(WD>=180)*(WD<=270),na.rm=TRUE)/(sum(dt,na.rm=TRUE)-sum(lost_dt,na.rm=TRUE)),
              dp_ec = sum(dpt2*(WD>=225)*(WD<=315),na.rm=TRUE)/(sum(dt,na.rm=TRUE)-sum(lost_dt,na.rm=TRUE)),
              dp_sec = (sum(dpt2*(WD>=270)*(WD<=360)+dpt2*(WD==0),na.rm=TRUE))/
                (sum(dt,na.rm=TRUE)-sum(lost_dt,na.rm=TRUE)),
              rrdp_n = sum(dpt2_n_comp, na.rm=TRUE)/(sum(dt,na.rm=TRUE)-sum(lost_dt,na.rm=TRUE)),
              rrdp_e = sum(dpt2_e_comp, na.rm=TRUE)/(sum(dt,na.rm=TRUE)-sum(lost_dt,na.rm=TRUE)),
              nobs=n()) %>%
    mutate(kept_dt_su =  dt_su - lost_dt_su)
  return(tib)
}

#drop years with more than 175 (default) hours of lost time
drop_big_lost_dt_yrs<- function(tib,dtmax=175){ 
  tib <- tib[tib$lost_dt_su<=dtmax,]
  return(tib)
}

#a function to group the data by years - used for decadal computations
group_by_yrs <- function(tib,breaks){ 
  tib <- tib %>%
    mutate(num_yrs = 1) %>%
    group_by(gr=cut(yr,breaks=breaks,right=FALSE)) %>%
    summarize_at(vars(dpt2_su:num_yrs),sum,na.rm=TRUE)
  return(tib)
}

#a function to group the data by years - used for decadal computations
group_by_yrs_vu <- function(tib,breaks){ 
  tib <- tib %>%
    mutate(num_yrs = 1) %>%
    group_by(gr=cut(yr,breaks=breaks,right=FALSE)) %>%
    summarize(dpt2_su = sum(dpt2_su*kept_dt_su,na.rm=TRUE)/(sum(kept_dt_su,na.rm=TRUE)),
              dp_n = sum(dp_n*kept_dt_su,na.rm=TRUE)/(sum(kept_dt_su,na.rm=TRUE)),
              dp_nne = sum(dp_nne*kept_dt_su,na.rm=TRUE)/(sum(kept_dt_su,na.rm=TRUE)),
              dp_ne = sum(dp_ne*kept_dt_su,na.rm=TRUE)/(sum(kept_dt_su,na.rm=TRUE)),
              dp_ene = sum(dp_ene*kept_dt_su,na.rm=TRUE)/(sum(kept_dt_su,na.rm=TRUE)),
              dp_e = sum(dp_e*kept_dt_su,na.rm=TRUE)/(sum(kept_dt_su,na.rm=TRUE)),
              dp_ese = sum(dp_ese*kept_dt_su,na.rm=TRUE)/(sum(kept_dt_su,na.rm=TRUE)),
              dp_se = sum(dp_se*kept_dt_su,na.rm=TRUE)/(sum(kept_dt_su,na.rm=TRUE)),
              dp_sse = sum(dp_sse*kept_dt_su,na.rm=TRUE)/(sum(kept_dt_su,na.rm=TRUE)),
              dp_s = sum(dp_s*kept_dt_su,na.rm=TRUE)/(sum(kept_dt_su,na.rm=TRUE)),
              dp_ssw = sum(dp_ssw*kept_dt_su,na.rm=TRUE)/(sum(kept_dt_su,na.rm=TRUE)),
              dp_sw = sum(dp_sw*kept_dt_su,na.rm=TRUE)/(sum(kept_dt_su,na.rm=TRUE)),
              dp_wsw = sum(dp_wsw*kept_dt_su,na.rm=TRUE)/(sum(kept_dt_su,na.rm=TRUE)),
              dp_w = sum(dp_w*kept_dt_su,na.rm=TRUE)/(sum(kept_dt_su,na.rm=TRUE)),
              dp_wnw = sum(dp_wnw*kept_dt_su,na.rm=TRUE)/(sum(kept_dt_su,na.rm=TRUE)),
              dp_nw = sum(dp_nw*kept_dt_su,na.rm=TRUE)/(sum(kept_dt_su,na.rm=TRUE)),
              dp_nnw = sum(dp_nnw*kept_dt_su,na.rm=TRUE)/(sum(kept_dt_su,na.rm=TRUE)),
              dp_sc = sum(dp_sc*kept_dt_su,na.rm=TRUE)/(sum(kept_dt_su,na.rm=TRUE)),
              dp_swc = sum(dp_swc*kept_dt_su,na.rm=TRUE)/(sum(kept_dt_su,na.rm=TRUE)),
              dp_wc = sum(dp_wc*kept_dt_su,na.rm=TRUE)/(sum(kept_dt_su,na.rm=TRUE)),
              dp_nwc = sum(dp_nwc*kept_dt_su,na.rm=TRUE)/(sum(kept_dt_su,na.rm=TRUE)),
              dp_nc = sum(dp_nc*kept_dt_su,na.rm=TRUE)/(sum(kept_dt_su,na.rm=TRUE)),
              dp_nec = sum(dp_nec*kept_dt_su,na.rm=TRUE)/(sum(kept_dt_su,na.rm=TRUE)),
              dp_ec = sum(dp_ec*kept_dt_su,na.rm=TRUE)/(sum(kept_dt_su,na.rm=TRUE)),
              dp_sec = sum(dp_sec*kept_dt_su,na.rm=TRUE)/(sum(kept_dt_su,na.rm=TRUE)),
              rrdp_n = sum(rrdp_n*kept_dt_su,na.rm=TRUE)/(sum(kept_dt_su,na.rm=TRUE)),
              rrdp_e = sum(rrdp_e*kept_dt_su,na.rm=TRUE)/(sum(kept_dt_su,na.rm=TRUE)),
              num_yrs = sum(num_yrs)) %>%
    mutate(rdp = sqrt(rrdp_n^2 + rrdp_e^2),
           rdd = (atan2(rrdp_e,rrdp_n)*180/pi+180) %% 360  )
  return(tib)
}

# dp_plot <- function(dp,yscale=30){ #plot fryberger diagrams
#   dir <- seq(0,360-22.5,by=22.5)
#   ndp <- dp*cos(dir*pi/180) #north components of dp
#   edp <- dp*sin(dir*pi/180) #east components of dp
#   rndp <- sum(ndp) #north component of resultant dp
#   redp <- sum(edp) #east component of resultant dp
#   rdp <- sqrt(rndp^2+redp^2) #rdp magnitude
#   #rdp direction (rotated 180 to point where the sand is going)
#   rdir <- (atan2(redp,rndp)*180/pi+180) %% 360 
#   
#   dp_df <- structure(list(dir = dir, dp = dp),.Names=c("dir", "dp"),
#                      class="data.frame",row.names=c(NA,-16) )
#   
#   rdp_df <- structure(list(dir = rdir, dp = rdp),.Names=c("dir", "dp"),
#                       class="data.frame",row.names=c(17L) )
#   
#   #normalize by RDP - details are weird. Set to match scaling of an earlier plot
#   if(yscale=="RDP"){yscale <- 2.7e7*rdp_df[1,2]/13075309} 
#   
#   #Set ylim to be as large as needed for largest dp in set of diagrams being 
#   # displayed together. Then we get consistent scaling
#   base <- ggplot(dp_df, aes(x=dir,y=dp))
#   p <- base + coord_polar() + ylim(0,yscale) +
#     scale_x_continuous(limits=c(0,360),breaks=dir)
#   q <- p + geom_segment(data = dp_df , aes(y=0,xend=dir,yend=dp),col="black") +
#     geom_segment(data = rdp_df , aes(y=0,xend=dir,yend=dp),
#                  arrow=arrow(length=unit(0.3,"cm")),col="grey",size=1) +
#     theme(axis.line=element_blank(),axis.text.x=element_blank(),
#           axis.text.y=element_blank(),axis.ticks=element_blank(),
#           axis.title.x=element_blank(),
#           axis.title.y=element_blank(),legend.position="none",
#           panel.background=element_blank(),panel.border=element_blank(),
#           panel.grid.major=element_blank(),
#           panel.grid.minor=element_blank(),plot.background=element_blank()) 
#   #print(q)
#   return(q)
# }

#expects vector of 16 directional drift potentials starting at N and ending at NNW, rdp, and rdd 
dp_plot <- function(dp,rdp,rdd,yscale=30){ #plot fryberger diagrams
  dir <- seq(0,360-22.5,by=22.5)
  
  dp_df <- structure(list(dir = dir, dp = dp),.Names=c("dir", "dp"),
                     class="data.frame",row.names=c(NA,-16) )
  
  rdp_df <- structure(list(dir = rdd, dp = rdp),.Names=c("dir", "dp"),
                      class="data.frame",row.names=c(17L) )
  
  #normalize by RDP - details are weird. Set to match scaling of an earlier plot
  if(yscale=="RDP"){yscale <- 2.7e7*rdp_df[1,2]/13075309} 
  
  #Set ylim to be as large as needed for largest dp in set of diagrams being 
  # displayed together. Then we get consistent scaling
  base <- ggplot(dp_df, aes(x=dir,y=dp))
  p <- base + coord_polar() + ylim(0,yscale) +
    scale_x_continuous(limits=c(0,360),breaks=dir)
  q <- p + geom_segment(data = dp_df , aes(y=0,xend=dir,yend=dp),col="black") +
    geom_segment(data = rdp_df , aes(y=0,xend=dir,yend=dp),
                 arrow=arrow(length=unit(0.3,"cm")),col="grey",size=1) +
    theme(axis.line=element_blank(),axis.text.x=element_blank(),
          axis.text.y=element_blank(),axis.ticks=element_blank(),
          axis.title.x=element_blank(),
          axis.title.y=element_blank(),legend.position="none",
          panel.background=element_blank(),panel.border=element_blank(),
          panel.grid.major=element_blank(),
          panel.grid.minor=element_blank(),plot.background=element_blank()) 
  print(q)
  return(q)
}

# #plot and maybe save fryberger diagram for a particular row of the data
# dp_plot_yrs <- function(tib,row,yscale=30,fname=NULL){ 
#   st <- which(names(tib)=="dp_n")
#   en <- which(names(tib)=="dp_nnw")
#   dp <- as.numeric(tib[row,st:en])
#   #print(dp)
#   #print(str(dp))
#   dp_plot(dp,yscale=yscale)
#   if(!is.null(fname)){
#     ggsave(fname,height = 3, width=3, units= "in")
#   }
# }

#plot and maybe save fryberger diagram for a particular row of the data
dp_plot_yrs <- function(tib,row,yscale=30,fname=NULL){ 
  st <- which(names(tib)=="dp_n")
  en <- which(names(tib)=="dp_nnw")
  dp <- as.numeric(tib[row,st:en])
  rdd <- as.numeric(tib[row,"rdd"])
  rdp <- as.numeric(tib[row,"rdp"])
  #print(dp)
  #print(str(dp))
  dp_plot(dp,rdp,rdd,yscale=yscale)
  if(!is.null(fname)){
    ggsave(fname,height = 3, width=3, units= "in")
  }
}

#Compute directional dps as proportion of total dp
compute_rel_dp <- function(tib){ 
  levs <- rev(c("e","ese","se","sse","s","ssw","sw","wsw","w","wnw",
                "nw","nnw","n","nne","ne","ene"))
  tib <- tib %>% mutate_at(vars(dp_n:dp_nnw),funs(./dpt2_su)) %>%
    select(station:dp_nnw) %>% select(-(dp_m:lost_dt_su)) %>%
    pivot_longer(!c(station,yr),names_to="dir",values_to="dp") %>%
    mutate(dir=factor(gsub(".*_","",dir),levels=levs))
  return(tib)
}

#regression analyses for dp vs yr
lm_dp_vs_yr <- function(tib,start_yr=1961,with_plots=TRUE){ 
  lm_dp_yr <- tib %>% select(yr,dpt2_su) %>% filter(yr>=start_yr) %>%
    lm(dpt2_su ~ yr, data = .) %>% summary()
  if(with_plots){
    tib <- tib %>% select(yr,dpt2_su) %>% filter(yr>=start_yr)
    p <- ggplot(data=tib, mapping=aes(x=yr,y=dpt2_su)) + geom_point() +
      ylab('drift potential') + xlab('year') + 
      geom_smooth(method='lm')
    print(p)
  }
  return(lm_dp_yr)
}

#proportion change 1961-2019 based on linear model
lm_pred_prop_change <- function(lmsummary){
  m <- lmsummary$coefficients[2,1] #slope
  b <- lmsummary$coefficients[1,1] #intercept
  prop_change <- m*(2019-1961)/(b+m*1961)
  return(prop_change)
}
```

Import dp data for each station, drop rows with NA wind speeds, compute time intervals, compute drift potentials (q*dt, calling it dpt, dp is drift potential per time). Drop partial years at the beginning and end of the data. 1988 is missing from the data at BEH. Make a list of data frames to streamline what would otherwise be repetitive function calls.

```{r}
beh <- readRDS('clean_data/BEH_19730101_20201207_dp.rds') %>% compute_dp_stuff() %>% 
  drop_yrs(c(1973,1988,2020))
biv <- readRDS('clean_data/KBIV_19961231_20201203_dp.rds') %>% compute_dp_stuff() %>% 
  drop_yrs(c(1996,2020))
grb <- readRDS('clean_data/GRB_19490901_20201207_dp.rds') %>% compute_dp_stuff() %>% 
  drop_yrs(c(1949,2020))
mdw <- readRDS('clean_data/MDW_19480101_20201207_dp.rds') %>% compute_dp_stuff() %>% 
  drop_yrs(c(2020))
mke <- readRDS('clean_data/MKE_19471231_20201207_dp.rds') %>% compute_dp_stuff() %>% 
  drop_yrs(c(1947,2020))
mkg <- readRDS('clean_data/KMKG_19480101_20201204_dp.rds') %>% compute_dp_stuff() %>% 
  drop_yrs(c(2020))
tvc <- readRDS('clean_data/TVC_19481201_20201207_dp.rds') %>% compute_dp_stuff() %>% 
  drop_yrs(c(1948,2020))

station_list <- list(beh=beh,biv=biv,grb=grb,mdw=mdw,mke=mke,mkg=mkg,tvc=tvc)
```

Compute yearly drift potentials, including by direction. The dpt2 variable computes drift potentials for at most a 3 hour duration. If the gap between reports is larger, then dpt2 is dp*3 (just the first 3 hours), and the rest of the duration is ignored (0 dp during the rest). The lost_dt variable measures how much time is lost. We remove years from the data where this adds up to more than 175 hours (2% of the year). In some cases, stations seemed to stop reporting at night (see, e.g., early BEH data). These years will not be represented in the data. In other cases there were large gaps in the data. These years are also removed.

```{r}
lost_dt_max <- 175 #hrs
station_list_yr <- station_list %>% map(compute_by_yr_stuff_directional_vu) %>%
  map(drop_big_lost_dt_yrs,dtmax=lost_dt_max) #as a list of tibbles
#as a single tibble (better for plotting)
station_yr <- station_list_yr %>% bind_rows(.id='station') 
station_yr %>% write_csv("clean_data/station_yr_vu.csv")
```

Now create decadal summaries. When doing decadal summaries this way, it is important to be mindful of missing years introduced in the previous step. This is especially important if looking at decadal dp totals, for example. To make comparisons of amounts of drift potential fair this should be divided by the number of years retained in the decade. The number of years retained is located in the num_yrs column. We do not do this sort of comparison. Decades are 1948-1957, 1958-1967,..., 2008-2017.

```{r}
breaks <- seq(1948,2018,by=10)
#as a list of tibbles
station_list_dec <- station_list_yr %>% map(group_by_yrs_vu,breaks=breaks) 
station_dec <- station_list_dec %>% bind_rows(.id='station') #as a single tibble

stat_dec_numyrs <- station_dec %>% select(station,gr,num_yrs)
print(stat_dec_numyrs)
```

Plot the total drift potential for each year and for each station.

```{r}

(p <- ggplot(data = station_yr, 
             aes(x = yr, y = dpt2_su, shape = station, color = station)) +
   geom_point() +
   geom_smooth(se = FALSE, size = 0.5) +
   scale_shape_manual(values=c(15,16,17,18,19,3,8)) +
   scale_color_viridis(discrete = TRUE, option = 'turbo') +
   theme_minimal() +
   ylab('Drift Potential (VU)') +
   xlab('Year'))

ggsave('outputs/dp_yearly.pdf')
```
Regression analysis for each station, using only years after 1960.

```{r}
lm_list <- station_list_yr %>% map(lm_dp_vs_yr,start_yr=1961,with_plots=TRUE)

#proportion change over 1961-2019 using linear model
prop_change_list <- lm_list %>% map(lm_pred_prop_change)
```

Just nicely formatted plots.

```{r}
station_dp_yr <- station_yr %>% 
  filter(yr > 1960) %>%
  rename(DP = dpt2_su, Year = yr, Station = station) %>%
  select(Year, DP, Station) %>%
  mutate(Station = recode(Station, "beh" = "BEH", "biv" = "BIV", 
                          "grb" = "GRB", "mdw" = "MDW", "mke" = "MKE", "mkg" = "MKG","tvc" = "TVC"  ))

ggplot(data = station_dp_yr, mapping = aes(x = Year, y = DP)) +
  geom_point() +
  facet_wrap(Station~., nrow = 4, scales = "free") +
  geom_smooth(method = 'lm', se = FALSE) +
  theme_minimal()

ggsave('outputs/DP_lm.pdf') 
```

Plot and save decadal Fryberger diagrams for 5 stations.

```{r}
fry_table_stations <- c("mkg","tvc","mke","grb","mdw")
#yscales <- rep(2.7e7,5)
yscales <- rep("RDP",5) #normalize to rdp
ros <- 1:7
yr_int_str <- c("1948_1957","1958_1967","1968_1977","1978_1987","1988_1997","1998_2007",
                "2008_2017")


for(i in seq_along(fry_table_stations)){ #loop over stations
  for(j in seq_along(ros)){ #loop over decades
    #plots are saved as pdfs. These are the file names.
    fname <- paste0("outputs/",fry_table_stations[i],"_",yr_int_str[j],".pdf") 
    #fname <- NULL # if not saving
    dp_plot_yrs(station_list_dec[[fry_table_stations[i]]],ros[j],yscale=yscales[i],
                fname=fname)
  }
}
```


```{r}
station_coast <- station_dec %>% 
  mutate(decade = recode_factor(gr, "[1.95e+03,1.96e+03)" = "1948-1957",
                                "[1.96e+03,1.97e+03)" = "1958-1967",
                                "[1.97e+03,1.98e+03)" = "1968-1977",
                                "[1.98e+03,1.99e+03)" = "1978-1987",
                                "[1.99e+03,2e+03)" = "1988-1997",
                                "[2e+03,2.01e+03)" = "1998-2007",
                                "[2.01e+03,2.02e+03)" = "2008-2017")) %>%
  select(station, decade, dp_sc, dp_nc, dp_ec, dp_wc, dpt2_su) %>%
  rename(dp = dpt2_su, sc = dp_sc, nc = dp_nc, ec = dp_ec, wc = dp_wc) %>%
  filter(!is.na(decade), station %>% is.element(c("tvc", "mkg", "mke", "mdw", "grb")))

station_coast_prop <- station_coast %>% 
  mutate(across(sc:wc, ~.x/dp)) %>%
  select(-dp) %>%
  pivot_longer(sc:wc, names_to = "coast", values_to = "prop") %>%
  mutate(station = recode(station, "grb" = "GRB", "mdw" = "MDW", "mke" = "MKE", "mkg" = "MKG","tvc" = "TVC" ))
```


```{r}
#to make things consistent with annual dp plot
station_colors <- viridis(7,option="turbo")[3:7]
station_symbols <- c(15,16,17,18,19,3,8)[3:7]

facet_labels <- c("Eastern DP", "Northern DP", "Southern DP", "Western DP")
names(facet_labels) <- c("ec", "nc", "sc", "wc")


ggplot(data = station_coast_prop,
       mapping = aes(x = decade, y = prop, shape = station, color = station, group = station)) +
  geom_point() +
  geom_path() +
  scale_color_manual(values=station_colors) +
  scale_shape_manual(values=station_symbols) +
  facet_wrap(coast ~ ., nrow = 2, ncol = 2,  scales = "free_y", labeller = labeller(coast = facet_labels)) +
  scale_x_discrete(guide = guide_axis(angle = 45)) +
  xlab("Decade") +
  ylab("Proportion of Total DP") +
  labs(color = "Station", shape = "Station") +
  theme_minimal()

ggsave('outputs/dp_coast_dec.pdf')  
```











